---
title: "Habitat1"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(mdsr)
library(mosaic)
library(ggridges)
library(RColorBrewer)
```

```{r}
species_general <- read_csv("species_general.csv")
```


Objective: Find and plot mean number of mammal oservations by month.

```{r}
#names(species_general)
#a <- mean(species_general19$mammals_no.wtracks~species_general19$Month, na.rm= TRUE)
#a.1 = a %>% 
#select("Jan", "Feb", "Mar", "April", "May", "June", "July", "Aug", "Sept", "Oct", "Nov", "Dec")
#a
mean_mdata = species_general19 %>%
  group_by(Month) %>%
  summarize(mean_mammals = mean(mammals_no.wtracks, na.rm = TRUE)) %>%
  slice(1:12) %>%
  mutate(Month = factor(Month, 
                        levels = c("Jan", "Feb", "Mar", "April", "May", "June",
                                   "July", "Aug", "Sept", "Oct", "Nov", "Dec")))
levels(mean_data$Month)
```

&For the birds....
```{r}
mean_bdata = species_general19 %>%
  group_by(Month) %>%
  summarize(mean_birds = mean(birds_no.wcalls, na.rm = TRUE)) %>%
  slice(1:12) %>%
  mutate(Month = factor(Month, 
                        levels = c("Jan", "Feb", "Mar", "April", "May", "June",
                                   "July", "Aug", "Sept", "Oct", "Nov", "Dec")))

```

&For the reptiles
```{r}
species_general19$repamp_spec = species_general19$`reptiles,amph_no.species` 
mean_radata = species_general19 %>%
  group_by(Month) %>%
  summarize(mean_reptiles_and_amph. = mean(repamp_spec, na.rm = TRUE)) %>%
  slice(1:12) %>%
  mutate(Month = factor(Month, 
                        levels = c("Jan", "Feb", "Mar", "April", "May", "June",
                                   "July", "Aug", "Sept", "Oct", "Nov", "Dec")))
```


&For the invertebrates
```{r}
mean_bugdata = species_general19 %>%
  group_by(Month) %>%
  summarize(mean_bugs = mean(bugs_no.species, na.rm = TRUE)) %>%
  slice(1:12) %>%
  mutate(Month = factor(Month, 
                        levels = c("Jan", "Feb", "Mar", "April", "May", "June",
                                   "July", "Aug", "Sept", "Oct", "Nov", "Dec")))
```

&For the fish
```{r}
mean_fdata = species_general19 %>% 
  group_by(Month) %>% 
  summarize(mean_fish= mean(fish, na.rm= TRUE)) %>% 
  slice(1:12) %>% 
  mutate(Month = factor(Month, 
                        levels = c("Jan", "Feb", "Mar", "April", "May", "June",
                                   "July", "Aug", "Sept", "Oct", "Nov", "Dec")))
```


Compiling the groups into Month Count
```{r}
month_count1 = mean_bdata %>% 
  inner_join(mean_bugdata, by= "Month")
month_count2= mean_fdata %>% 
  inner_join(mean_mdata, by= "Month")
month_count3 = month_count2 %>% 
  inner_join(mean_radata, by="Month")
month_count4 = month_count1 %>% 
  inner_join(month_count3, by= "Month")
#names(month_count)
month_count_tidy = month_count4 %>%
  gather(key = Animal_Group, value = Species_Richness, -Month)
month_count_tidy
```

Objective: Plot line graphs of number of animal observations per group of 2019 data
- I would love to make multiple density ridges or something on one plot
```{r}
ggplot(month_count_tidy, aes(x = Month, y = Species_Richness, fill = Animal_Group)) + 
  geom_col() + 
  labs(x = "Month", y = "Mean Number of Species Observed") + scale_fill_brewer(palette= "Set1") + theme_minimal() + ggtitle("Species Diversity by Animal Group and Month") + labs(caption = "Source: Mass Audubon Habitat Nature Center Citizen Science Observations")
```

Objective: Visualize when which bird species are spotted when during a calendar year

- needs to be fixed- I need to reduce the number of rows down to one per bird type per month

```{r}
species_birds3 = species_general19[, c(2, 25:48)]
species_birds_tidy2 = species_birds3 %>%
  gather(key = Species, value = Present, -Month) %>% 
  #gather(key = Month, value = Present, -Species)
View(species_birds_tidy2)
```


Objective: Identify which bird species is most frequently spotted

- Almost; if I run the whole chunk, the mutate() command will order the Bird types, but it adds an NA column which counts the number of less-common birds that I don't want.
```{r}
#names(species_general19)
species_birds = species_general19[, c(1, 25:48)]
species_birds_tidy = species_birds %>%
  gather(key = Species, value = Present, -Date)
#(species_birds_tidy)
species_birds2 = species_birds_tidy %>%
  filter(Present == 1) %>%
  count(Species, sort = TRUE) %>% 
  top_n(n= 5, wt= n) %>% 
  mutate(Species = factor(Species, 
                        levels = c("Bluejay", "Ducks", "robin", "Cardinals", "owl", "woodpecker")))
#species_birds2
ggplot(species_birds2, aes(y = Species, x = n)) +
  geom_segment(aes(x = 0, y = Species, xend = n, yend = Species), color = "red", size = 0.75) +
  geom_point(size = 3) + labs(x= "Number of times bird type was observed", y= "Bird type")+ ggtitle("Most Frequently Observed Birds")
```

Objective: What are the best conditions to see a turtle?
- why isn't the correlation matrix generating?

```{r}
#View(species_general19)
cor4turtles = species_general19 %>% 
  select(Temperature_F, Group_size, Duration_hrs, Month, Time, Wind, cloud_cover, Turtles)
#cor4turtles_untidy = cor4turtles %>% 
  #spread(key= Time, value =AM)
#cor4turtles
numbers_only_turtles <- lm(Turtles~Temperature_F+Group_size+ Duration_hrs, data=cor4turtles)
summary(numbers_only_turtles)
cor(cor4turtles)
```

According to the above statistics, temperature is a very good predictor of the presence or absence of turtles. However, none of the other numeric predictors are a good secondary predictor.

Objective: How many coyotes/coyote tracks were spotted?

```{r}
count(species_general19$coyote)
2/75
```
The answer is two/ 3% of observers.

Invasive v. Non-invasive birds

```{r}

  
```


Research Question: What is the best predictor of insect presence?
- I am wondering if people who spend longer making observations, larger groups, etc. are more likely to identify more insects, or are some people more in-tune with these parts of nature? Or is it weather?
```{r}

```



Research Question: Does time of day impact anything? Is there a "best" time to see wildlife at Habitat?
```{r}

```

Questions to answer:

When were the more rare species spotted in 2018-2019?
Under which conditions (time of year, day, weather) are visitors most likely to see turtles?
Is there anything related to whether birds are invasive or not?
What sorts of people write down insect observations? Is it school groups, people who spend longer making wildlife observations, etc.?
Is there a "best" time of day to see wildlife at Habitat?

